Uncovering underlying high-level musical content in the time domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn deep audio embeddings with applications to MIR downstream tasks

### 2023-06-13 evaluation/evaluation.tex
Updates from Overleaf

While the usefulness of deep audio embeddings can be evaluated in countless downstream tasks \cite{Li2023MERT:Training, Kim2020OneStrategies}, we have chosen music boundary detection, a subset of Music Structure Analysis (MSA) for its popularity, complexity, and product-compelling nature. This interdisciplinary field aims to understand the structure of music \cite{Nieto2020Audio-BasedApplications}. Due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges like boundary placement ambiguity and similarity quantification \cite{NietoPerceptualMusic}.

@@

All the evaluation metrics have been computed using the mir-eval package \cite{RaffelMir_eval:METRICS}.

>>>>

While the usefulness of deep audio embeddings can be evaluated in countless downstream tasks \cite{Li2023MERT:Training, Kim2020OneStrategies}, we have chosen music boundary detection, a subset of Music Structure Analysis (MSA) for its popularity \cite{Smith2013ATask}, complexity, and product-compelling nature. This interdisciplinary field aims to understand the structure of music \cite{Nieto2020Audio-BasedApplications}. Due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges like boundary placement ambiguity and similarity quantification \cite{NietoPerceptualMusic}.

@@

### 2023-06-13 figures/neural networks/activation.tex
Updates from Overleaf

\begin{figure}
    \centering
    \scalebox{0.9}{
% NEURAL NETWORK activation
% https://www.youtube.com/watch?v=aircAruvnKk&list=PLZHQObOWTQDNU6R1_67000Dx_ZCJB-3pi&index=1
\begin{tikzpicture}[x=2.7cm,y=1.6cm]
\centering
  \message{^^JNeural network activation}
  \def\NI{5} % number of nodes in input layers
  \def\NO{4} % number of nodes in output layers
  \def\yshift{0.4} % shift last node for dots

@@

\end{tikzpicture}
}
\caption[Activation function.]{\small{The input layer with \textit{n} nodes is denoted as $a_i^{(0)}$ for $i \in \{1, \ldots, n\}$. The output layer with \textit{m} nodes is $a_j^{(1)}$ for $j \in \{1, \ldots, m\}$. Connections between these layers are governed by the weight matrix $\mathbf{W}^{(0)} \in \mathbb{R}^{m\times n}$ and bias vector $\mathbf{b}^{(0)} \in \mathbb{R}^{m}$, with elements $w_{j,i}$ and $b_j^{(0)}$ respectively. The activation function, denoted as $\sigma$, operates element-wise on the sum of the product of the weight matrix, input vector, and bias vector.}}

>>>>

\begin{figure}
    \centering
    \scalebox{0.8}{
% NEURAL NETWORK activation
% https://www.youtube.com/watch?v=aircAruvnKk&list=PLZHQObOWTQDNU6R1_67000Dx_ZCJB-3pi&index=1
\begin{tikzpicture}[x=2.7cm,y=1.6cm]
\centering
  \message{^^JNeural network activation}
  \def\NI{5} % number of nodes in input layers
  \def\NO{4} % number of nodes in output layers
  \def\yshift{0.4} % shift last node for dots

@@

\end{tikzpicture}
}
\caption[Activation function \cite{tikz}.]{\small{The input layer with \textit{n} nodes is denoted as $a_i^{(0)}$ for $i \in \{1, \ldots, n\}$. The output layer with \textit{m} nodes is $a_j^{(1)}$ for $j \in \{1, \ldots, m\}$. Connections between these layers are governed by the weight matrix $\mathbf{W}^{(0)} \in \mathbb{R}^{m\times n}$ and bias vector $\mathbf{b}^{(0)} \in \mathbb{R}^{m}$, with elements $w_{j,i}$ and $b_j^{(0)}$ respectively. The activation function $\sigma$ operates element-wise on the sum of the product of the weight matrix, input vector, and bias vector.}}

### 2023-06-13 introduction/Subsections/Previous work.tex
Updates from Overleaf

Previous studies have highlighted the significance of selecting suitable audio features to distinguish and label unique music segments in music segmentation tasks. However, acquiring music segment annotation data for learning feature transformations presents a challenge due to its labor-intensive and expensive nature. Research has leveraged unsupervised deep learning to mitigate this, using readily available music audio. This methodology has significantly enhanced the efficacy of segmentation algorithms, yielding state-of-the-art results. Furthermore, strategies that recognize the hierarchical nature of the structural organization employing two-level annotations of primary and secondary boundaries have further augmented boundary detection performance \cite{unsupervisedlearndeepfeat, GrillMUSICANNOTATIONS, Hernandez-Olivan2021MusicFeatures, SalamonDeepSegmentation}.

Apart from music structure applications, recent advances in neural networks have facilitated the generation of low-dimensional numerical representations that encapsulate key musical attributes. This development simplifies the computation of various task-specific elements. Such tasks encompass:

>>>>

Prior research underscores the importance of choosing appropriate audio features to differentiate and label distinct music segments for music segmentation tasks. The hurdle lies in procuring annotated data for learning these feature transformations, which can be laborious and costly. To overcome this, scholars have turned to unsupervised deep learning, utilizing accessible music audio \cite{unsupervisedlearndeepfeat, GrillMUSICANNOTATIONS}. This approach has markedly enhanced the generalization ability of machine-listening models in downstream applications, with segmentation algorithms reaping benefits in the form of state-of-the-art outcomes \cite{Hernandez-Olivan2021MusicFeatures, SalamonDeepSegmentation, teacher, Li2023MERT:Training}.

Beyond music structure applications, modern advancements in neural networks have facilitated the creation of condensed numerical representations that capture critical musical characteristics. This progress has streamlined the calculation process for various task-specific elements. Such tasks include:

### 2023-06-13 methods/subsections/implementation.tex
Updates from Overleaf

Contrastive Learning of Musical Representations (CLMR) is a method that learns valuable, discriminative music representations without explicit labels by contrasting positive augmentations of a musical piece against negative ones \cite{CLMR2021}. CLMR falls under a subset of ML called self-supervised learning (SSL) \cite{Balestriero2023ALearning}, where models learn independently from unlabeled data to create their supervisory signals \cite{audioselfsupsurvey}. This method resembles how humans learn from observations and interactions, turning unsupervised problems into supervised ones by auto-generating labels. Benefits of SSL include reduced dependence on labeled data, which contributes to more robust and generalizable data representations.

The methodology of CLMR involves generating different augmentations of the same musical piece to serve as positive pairs while using other pieces as antagonistic pairs. The training process, which minimizes a contrastive loss function, encourages the model to produce similar representations for positive and dissimilar ones for antagonistic pairs.

@@

There are several loss functions used in tasks like feature learning, similarity learning, and dimensionality reduction.

Functions such Mean Squared Error (MSE) Loss \cite{Hodson2021MeanDeconstructed}, Cross-Entropy Loss \cite{Zhang2018GeneralizedLabels}, Contrastive Loss \cite{supercontrast} or Triplet Loss are used when training models to understand the difference between various classes or categories.

Schroff, F., Kalenichenko, D., and Philbin, J. at Google initially proposed and used triplet loss to learn face recognition of the same person at different poses and angles. \cite{Schroff2015FaceNet:Clustering}

The triplet loss function guides the learning process. It seeks to minimize the distance between the anchor and positive instances while maximizing the distance between the anchor and the negative instances. A margin parameter is included in the loss function to ensure a minimum separation between positive and negative instances in the embedding space.

>>>>

Our implementation is a slight tweak to the \textit{contrastive learning of lusical representations} (CLMR), a method that learns valuable, discriminative music representations without explicit labels by contrasting positive augmentations of a musical piece against negative ones \cite{CLMR2021}. CLMR falls under a subset of ML called self-supervised learning (SSL) \cite{Balestriero2023ALearning}, where models learn independently from unlabeled data to create their supervisory signals \cite{audioselfsupsurvey}. This method resembles how humans learn from observations and interactions, turning unsupervised problems into supervised ones by auto-generating labels. Benefits of SSL include reduced dependence on labeled data, which contributes to more robust and generalizable data representations.

As stated above, the methodology of CLMR involves generating different augmentations of the same musical piece to serve as positive pairs while using other pieces as antagonistic pairs. The training process, which minimizes a triplet loss function, encourages the model to produce similar representations for positive and dissimilar ones for antagonistic pairs.

@@

Contrastive loss and triplet loss are two prominent loss functions employed in the fields of feature learning, similarity learning, and dimensionality reduction.

Schroff, F., Kalenichenko, D., and Philbin, J. from Google first proposed and applied triplet loss for the learning of facial recognition, catering to varied poses and angles of the same individual \cite{Schroff2015FaceNet:Clustering}.

Triplet loss function directs the learning process by endeavoring to minimize the distance between the anchor and positive instances and maximize the distance between the anchor and negative instances. Contrastive loss, on the other hand, works on pairs of instances \cite{supercontrast}. The inclusion of a margin parameter in the loss function guarantees a minimum separation between positive and negative instances in the embedding space. Given the nature of our training needs, we propose to train by using the former.

### 2023-06-13 results/subsections/tables.tex
Updates from Overleaf

SALAMI (Structural Analysis of Large Amounts of Music Information) is a project aiming to carry out extensive structural analyses on a wide variety of music \cite{Smith2011DESIGNANNOTATIONS}. Instead of traditional forms, the authors segment pieces into sections, combining perceptual, functional, and transcription analyses. Despite limitations, this approach offers a holistic and nuanced perspective on music structure.

METRICS

\begin{itemize}
    \item \textbf{Precision}: Precision measures the proportion of correctly detected boundaries out of the total estimated boundaries. It quantifies the accuracy of the boundary detection algorithm.

    \[
    \text{Precision} = \frac{\text{True Positives}}{\text{True Positives} + \text{False Positives}}
    \]

    High precision indicates a low rate of false positives, meaning the algorithm correctly identifies boundaries without many incorrect detections.

    \item \textbf{Recall}: Recall measures the proportion of correctly detected boundaries out of the total reference boundaries. It quantifies the completeness or sensitivity of the boundary detection algorithm.

    \[
    \text{Recall} = \frac{\text{True Positives}}{\text{True Positives} + \text{False Negatives}}
    \]

    High recall indicates a low rate of false negatives, meaning that the algorithm successfully detects most of the true boundaries without missing many.

    \item \textbf{F-measure}: The F-measure is a widely used boundary detection metric computed by comparing the predicted and ground truth boundaries. This metric, ranging from 0 to 1, represents the harmonic mean of precision and recall, providing a balanced view of under-segmentation and over-segmentation. Given the potential inaccuracies in human annotations and prediction errors, the F-measure tolerates minor discrepancies between predicted and ground truth boundaries. This tolerance level can be adjusted, enabling the metric to consider a predicted boundary as correct if it's within the set tolerance window of a ground truth boundary \cite{Turnbull2007ABOOSTING}.

    \[
    \text{F-measure} = \frac{2 \times \text{Precision} \times \text{Recall}}{\text{Precision} + \text{Recall}}
    \]
\end{itemize}

@@

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{4.5cm}XXXXX}
\toprule
\thead{\centering\textbf{Authors [Ref], Year}} & \thead{\centering\textbf{Input}} & \thead{\centering\textbf{Method}} & \thead{\centering\textbf{F-measure}} \\
\midrule
\addlinespace
Kaiser et al. \cite{27}, 2012 & SSM & Novelty measure  & 0.286 \\
\addlinespace
McFee \& Ellis \cite{20}, 2013 & MLS & Fisher’s Linear Discriminant  & 0.317 \\
\addlinespace
Nieto \& Bello \cite{28}, 2014 & MFCCs, chromas & Checkerboard-like kernel  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2015 & Timbre-type histograms & HMM  & 0.213 \\
\addlinespace
Nieto \cite{30}, 2016 & Constant-Q Transform Spectrogram & Linear Discriminant Analysis  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2017 & Timbre-type histograms & HMM  & 0.212 \\
\addlinespace
Turnbull et al. \cite{Turnbull2007ABOOSTING}, 2007 & MFCCs, chromas, spectrogram & Boosted Decision Stump  & 0.378 \\
\addlinespace
Sargent et al. \cite{34}, 2011 & MFCCs, chromas & Viterbi  & 0.356 \\
\addlinespace
Ullrich et. al \cite{22}, 2014 & MLS & CNN  & 0.465 \\
\addlinespace
Grill \& Schlüter \cite{4}, 2015 & MLS, SSLMs & CNN  & 0.523 \\
\addlinespace
Grill \& Schlüter \cite{GrillMUSICANNOTATIONS}, 2015 & MLS, PCPs, SSLMs & CNN  & 0.508 \\
\addlinespace
Hadria \& Peeterss \cite{35}, 2017 & MLS, SSLMs & CNN  & 0.291 \\
\bottomrule
\end{tabularx}
\caption[Baseline. State-of-the-art table.]{\small{Previous studies' boundary detection f-measure results using unsupervised methods for a 0.5s time-window tolerance. Only the top-performing algorithm for each year on the SALAMI dataset is displayed. Original source \cite{Hernandez-Olivan2021MusicFeatures}}}
\label{tab:my_label}
\end{table}

>>>>

The evaluation results have been obtained using the standard and most commonly used MIR's tools and frameworks such as MSAF \cite{NietoMSAF:FRAMEWORK} for the segmentation algorithms, the SALAMI dataset \cite{Smith2011DESIGNANNOTATIONS} as evaluation ground truth and $mir_eval$ \cite{RaffelMir_eval:METRICS} for metric computation.

@@

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{4.5cm}XXXXX}
\toprule
\thead{\centering\textbf{Authors [Ref], Year}} & \thead{\centering\textbf{Input}} & \thead{\centering\textbf{Method}} & \thead{\centering\textbf{F-measure}} \\
\midrule
\addlinespace
Kaiser et al. \cite{27}, 2012 & SSM & Novelty measure  & 0.286 \\
\addlinespace
McFee \& Ellis \cite{20}, 2013 & MLS & Fisher’s Linear Discriminant  & 0.317 \\
\addlinespace
Nieto \& Bello \cite{28}, 2014 & MFCCs, chromas & Checkerboard-like kernel  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2015 & Timbre-type histograms & HMM  & 0.213 \\
\addlinespace
Nieto \cite{30}, 2016 & Constant-Q Transform Spectrogram & Linear Discriminant Analysis  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2017 & Timbre-type histograms & HMM  & 0.212 \\
\addlinespace
Turnbull et al. \cite{Turnbull2007ABOOSTING}, 2007 & MFCCs, chromas, spectrogram & Boosted Decision Stump  & 0.378 \\
\addlinespace
Sargent et al. \cite{34}, 2011 & MFCCs, chromas & Viterbi  & 0.356 \\
\addlinespace
Ullrich et. al \cite{22}, 2014 & MLS & CNN  & 0.465 \\
\addlinespace
Grill \& Schlüter \cite{4}, 2015 & MLS, SSLMs & CNN  & 0.523 \\
\addlinespace
Grill \& Schlüter \cite{GrillMUSICANNOTATIONS}, 2015 & MLS, PCPs, SSLMs & CNN  & 0.508 \\
\addlinespace
Hadria \& Peeterss \cite{35}, 2017 & MLS, SSLMs & CNN  & 0.291 \\
\bottomrule
\end{tabularx}
\caption[Baseline. State-of-the-art table.]{\small{Previous studies' boundary detection f-measure results using unsupervised methods for a 0.5s time-window tolerance. Only the top-performing algorithm for each year on the SALAMI dataset is displayed. Original source \cite{Hernandez-Olivan2021MusicFeatures}}}
\label{tab:comparison_table}
\end{table}
